Edge effects in Alberta

### 2021-07-30 Manuscript/manuscript.Rmd
Updated model to filter extreme data

- Yield data is collected in rectangles of the same length as the data interval (distance = combine ground speed $\times$ interval, typically 1 second) and the same width as the combine header (5-7 m)
  - We extracted the size of each polygon (m$^2$), dry yield (tonnes), and the spatial location, and the sequence of collection (1 - end of harvest)
  - Because of the large number of yield rectangles per field (30-800 thousand), we used the centroid of each polygon as its location, treating areal data as point data
  - Seeding and application rates were constant across fields, so we did not consider inputs in our analysis
  - We used dry yield (tonnes of seed/hectare minus recorded crop moisture) as our measure of crop yield
  - Precision yield data can be highly variable and is prone to extreme outliers (both positive and negative), especially when combine ground speeds are low and at the beginning of rows
  - Therefore, we filtered data that lay outside the 95th percentiles of yield within each field, values less than or equal to zero, and values outside of the 95th percentiles of ground speed (high: 
  - Due to the large number of data at each field, we sub-sampled to 50,000 data points per field to reduce computation time

@@

- In addition to modeling edge effects (our variable of interest), this model also accounts for a) differences in combine speed, b) within-field spatial variation not related to edges, and d) shifts in combine accuracy during harvest
  - Spatial or temporal variation is typically modelled using a Gaussian Process Model (e.g. Kriging, SPDE Approximations) but this was computationally unfeasible with 50,000 data points per field
  - Penalized splines offer a compromise, as they account for nonlinear ``wiggly" relationships in the same way as Gaussian processes but using a reduced number of dimensions (as well as computation time)
  - The number of basis dimensions was checked with the _gam.check_ function from _mgcv_
  - The relationship between polygon size (i.e. ground speed) and yield was modeled with a log-linear relationship with a single slope term, as this closely matched smoothed versions
  - All models were fit in _R_ using the _mgcv_ library [version `r packageVersion('mgcv')`, @wood2017], and figures were created with _ggplot2_ and _ggpubr_ [versions `r packageVersion('ggplot2')`, @wickham2016; and `r packageVersion('ggpubr')`, @kassambara2020].

@@

>>>>

- Yield data is collected in rectangles of the same length as the data interval (distance = combine ground speed $\times$ interval, typically 1 second) and the same width as the combine header (5-7 m, adjusted to account for overlap with adjacent swaths)
  - We extracted the size of each polygon (m$^2$), dry yield (tonnes), and the spatial location, and the sequence of collection (1 - end of harvest)
  - Because of the large number of yield rectangles per field (30-800 thousand), we used the centroid of each polygon as its location, treating areal data as point data
  - Seeding and application rates were constant across fields, so we did not consider inputs in our analysis
  - We used dry yield (tonnes of seed/hectare minus recorded crop moisture) as our measure of crop yield
  - Precision yield data can be highly variable and is prone to extreme outliers (both positive and negative), especially when combine ground speeds are low and at the beginning of rows
  - Therefore, we filtered data that lay outside the 95th percentiles of yield within each field, values less than or equal to zero, and values outside of the 95th percentiles of ground speed (high: 
  - Due to the large number of data at each field, we sub-sampled to 50,000 data points per field to reduce computation time

@@

- In addition to modeling edge effects (our variable of interest), this model also accounts for a) differences in combine speed, b) within-field spatial variation not related to edges, and d) shifts in combine accuracy during harvest
  - Spatial or temporal variation can be modelled using a Gaussian Process Model (e.g. Kriging, SPDE Approximations) but this was computationally unfeasible with 50,000 data points per field
  - Penalized splines offer a compromise, as they account for nonlinear ``wiggly" relationships in the same way as Gaussian processes but using a reduced number of dimensions (as well as computation time)
  - The number of basis dimensions was checked with the _gam.check_ function from _mgcv_
  - The relationship between polygon size (i.e. ground speed) and yield was modeled with a log-linear relationship with a single slope term, as this closely matched smoothed versions
  - All models were fit in _R_ using the _mgcv_ library [version `r packageVersion('mgcv')`, @wood2017], and figures were created with _ggplot2_ and _ggpubr_ [versions `r packageVersion('ggplot2')`, @wickham2016; and `r packageVersion('ggpubr')`, @kassambara2020].

- Despite the use of spatial and temporal smoothers, the residuals in most of the models displayed a large amount of spatial and temporal autocorrelation
   - This is extremely common in precision yield data, but

@@

## Data

## Boundary effects
